Hector Permafrost

### 2020-02-06 analysis/paper/paper.Rmd
Move methods text around.

where $a$, $b$, and $c$ are model parameters.
The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:

@@

For globally-averaged permafrost, $a = 2.371$, $b = -0.676$, and $c = -3.685$ to most closely reproduce the rate of 0.172 K$^{-1}$ reported by [Kessler (2015)][kessler] over the range of 0.8 [^2] to 4 K above the pre-industrial baseline.

@@

sigmoid_cap <- paste(
  "Fraction of permafrost thaw as a function of change in",
  "global annual mean air temperature since pre-industrial (1750)."
)
pfa <- 2.371
pfb <- -0.676
pfc <- -3.685
b <- 4.078
kessler_m <- 0.172
kessler_b <- 0.8 * kessler_m + 1
plot(0, 0, type = "n", xlim = c(0, 8), ylim = c(0, 1),
     xlab = expression(Delta * T[air] ~ (K)),
     ylab = "Frac. permafrost remaining")
curve(1 - (1 + pfa * exp(pfb * x)) ^ pfc, 0, 10, col = 1, add = TRUE)
abline(a = kessler_b, b = -kessler_m, lty = "dashed")
abline(h = 1, lty = "dotted")
legend("topright", c("Hector", "Kessler (2015)"), lty = c("solid","dashed"),
       bg = "white")

@@

Initial permafrost C is set to 1035 Pg C.

>>>>

where $a$, $b$, and $c$ are model parameters whose values are described in the _Configuration_ section below.
The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:

@@

## Methane emissions

In the default configuration of Hector, natural CH~4~ emissions are prescribed at a constant value (300 Tg year^-1^) [@hartin_2015_simple].
In the new configuration, CH~4~ emissions are a fixed fraction of heterotrophic respiration.

Total heterotrophic respiration ($RH[i,t]$) for biome $i$ at time $t$ is the sum of heterotrophic respiration in detritus ($RH_d$) and soil ($RH_s$).

$$ RH[i,t] = RH_s[i,t] + RH_d[i,t] $$

Detritus and soil heterotrophic respiration are proportional to the sizes of their respective carbon pools ($C_d$ and $C_s$, respectively, both in Pg C),
with a rate that increases exponentially with temperature according to a biome-specific temperature sensitivity parameter ($Q_{10}$).
Detritus respiration increases with biome-specific air temperature change since pre-industrial ($T[i,t]$),
while soil respiraiton increases with the 200 year running mean of air temperature ($T_{200}[i,t]$) because soils warm slowly relative to the atmosphere.

\[ T_{air}[i,t] = \delta \, T_{air}[t] \]

\[ RH_d[i,t] = \frac{1}{4} C_d Q_{10}[i] ^ {T}[i,t] / 10} \]

\[ RH_s[i,t] = \frac{1}{50} C_s Q_{10}[i] ^ {T_{200}[i,t] / 10} \]

In the original formulation, all of the total heterotrophic respiration was assumed to be CO~2~ and natural CH~4~ emissions were entirely separate.
In the revised version here, total heterotrophic respiration is divided into CO~2~ ($RH_{CO_2}[i,t]$) and CH~4~ ($RH_{CH_4}[i,t]$) based on the biome-specific CH~4~ fraction parameter ($f_{CH_4}[i]$):

\[ RH_{CH_4}[i,t] = f_{CH_4}[i] RH[i,t] \]

\[ RH_{CO_2}[i,t] = (1 - f_{CH_4}[i]) RH[i,t] \]

Initial permafrost C is set to 1035 Pg C based on the estimate for near-surface (<3 m depth) permafrost by @hugelius_2014_estimated.

```{r params}
# Permafrost exponential model
pfa <- 2.371
pfb <- -0.676
pfc <- -3.685

# Kessler linear model
kessler_m <- 0.172
kessler_b <- 0.8 * kessler_m + 1
```

For globally-averaged permafrost, exponential parameters $a = `r pfa`$, $b = `r pfb`$, and $c = `r pfc`$ to most closely reproduce the rate of `r kessler_m` K^-1^ reported by [Kessler (2015)][kessler] over the range of 0.8 [^2] to 4 K above the pre-industrial baseline.

@@

sigmoid_cap <- paste(
  "Fraction of permafrost thaw as a function of change in",
  "global annual mean air temperature since pre-industrial (1750)."
)
plot(0, 0, type = "n", xlim = c(0, 8), ylim = c(0, 1),
     xlab = expression(Delta * T[air] ~ (K)),
     ylab = "Frac. permafrost remaining")
curve(1 - (1 + pfa * exp(pfb * x)) ^ pfc, 0, 10, col = 1, add = TRUE)
abline(a = kessler_b, b = -kessler_m, lty = "dashed")
abline(h = 1, lty = "dotted")
legend("topright", c("Hector", "Kessler (2015)"), lty = c("solid","dashed"),
       bg = "white")

@@
